---
title: "Chapter 6"
format: 
  html:
    embed-resources: true
author: Visruth
---

# 6.1

A standard idea of the product of two types is a pair. Pairs are commutative up to isomorphism.

```{r}
swap <- \(x, y) c(y, x)
```

swap is clearly swap's inverse.

The ordering can be thought of as just using a different format for storing data.

Nested pairs are equivalent to tuples since associated pairs are equal up to isomorphism.

Unit type is also the unit object for products.


```{r}
# NULL is inappropriate as unit, but good enough to get the point across
rho <- \(x, y) x
rho_inv <- \(x) c(x, "NULL")
```

Set is a monoidal category, i.e. it is a category which is also a monoid since you can multiply objects.

> It’s often preferable to give names to components. A product type with named fields is called a record in Haskell, and a struct in C. 

# 6.2

Some Haskell stuff

```{r}
library(S7)

Element <- new_class("Element", properties = list(name = class_character, symbol = class_character, atomicNumber = class_numeric))
helium <- Element("Helium", "He", 2)

helium@name
helium@symbol
helium@atomicNumber
```

Can easily use S7 classes as records.

# 6.3

Sum types are like Unions.

Lots of details on the differences between Haskell and C++ wrt to unions.

> In terms of types, we have Either as the monoidal operator and Void, the uninhabited type, as its neutral element.

Haskell data types are immutable which allows you to deconstruct them well after construction through pattern matching and a re-use of constructors.

# 6.4

We've constructed two commutative monoids using the type system. Sum types have Void, and product types unit type as the "neutral" element. Void ~ 0, unit ~ 1.

Intertwined monoids like these are called semi-rings. A full ring would need subtraction; we don't have subtraction of types. Semi-rings can be called rigs, since they are rings without negation. Ha ha.

Refer to textbook for very cool look into structure of a list--an algebraic perspective. We have an algebra on types, which is why they're called algebraic data types.

> Logical and & or also form a semiring, and it too can be mapped into type theory: 

See text for table, but it's self explanatory. Void is FALSE, unit is TRUE, A or B is Either A B, A and B is a tuple (A, B).

# 6.5

1.

Take maybes to eithers by giving back right value if just value exists, or left unit it doesn't. Simple to invert this too.

2.
The actual question is trivial, will answer using S7 which doesn't map completely.

```{r}
library(S7)

Shape <- new_class("Shape", abstract = TRUE)
Circle <- new_class("Circle", Shape, properties = list(radius = class_numeric))
Rect <- new_class("Rect", Shape, properties = list(width = class_numeric, height = class_numeric))

Area <- new_generic("Area", "x")
method(Area, Circle) <- \(x) pi * x@radius^2
method(Area, Rect) <- \(x) x@width * x@height

circ <- Circle(4 / (sqrt(pi)))
rect <- Rect(10, 4)

Area(circ)
Area(rect)
```

3.

```{r}
Circumference <- new_generic("Area", "x")
method(Circumference, Circle) <- \(x) 2 * pi * x@radius
method(Circumference, Rect) <- \(x) 2 * (x@width + x@height)

Circumference(circ)
Circumference(rect)
```

None of the original code.

4.

```{r}
Square <- new_class("Square", Rect, constructor = function(side) new_object(Rect, width = side, height = side))
square <- Square(5)

Area(square)
Circumference(square)
```

R can be quite functional, so again not much to be done. It's certainly not as clean as a Haskell implementation but clearly its very easy to extend these shapes using S7.

5.

a + a = 2 * a

a + a is Either a a = Left a | Right a

2 is data Bool = True | False 

2 * a then is (Bool, a)

So we need to find an iso from Either a a to (Bool, a). Take Either to Pair by giving the Bool in Pair TRUE if Left a, else FALSE. The inverse would then be mapping (TRUE, a) to Right a and the FALSE to Left a.

